---
title: "GHG processor"
author: "Bond-Lamberty / Wilson"
title-block-banner: true
params:
  html_outfile: "ghg-processor.html"
  data_folder: "test/"
  output_folder: "output/"
  metadata_folder: "test/"
  logfile: ""
date: now
date-format: "YYYY-MM-DD HH:mm:ssZ"
format: 
  html:
    code-fold: true
editor: visual
---

## Initializing

```{r}
suppressPackageStartupMessages({
  library(readr)
  library(lubridate)
  library(ggplot2)
  library(readxl)
})
theme_set(theme_bw())

data_files <- list.files(params$data_folder, 
                         pattern = "*txt$", 
                         full.names = TRUE, recursive = TRUE)
metadata_files <- list.files(params$metadata_folder, 
                             pattern = "*xlsx$", 
                             full.names = TRUE, recursive = TRUE)
```

I see `r length(data_files)` data files to process.

I see `r length(metadata_files)` metadata files to process.

Writing output to `r normalizePath(params$output_folder)`.

HTML outfile is `r params$html_outfile`.

Logfile is `r params$logfile`.

Working directory is `r getwd()`.

## Read in GHG data

```{r}
errors <- 0

read_ghg_data <- function(fn) {
  basefn <- basename(fn)
  message(Sys.time(), " Processing ", basefn)
  
  dat_raw <- readLines(fn)
  
  # These files have five header lines, the names of the columns in line 6,
  # and then the column units in line 7
  dat <- try({
    readr::read_table(I(dat_raw[-c(1,2,3,4,5,7)]), na = "nan")
  })
  
  if(is.data.frame(dat)) {
    # Save the serial number in case we want to look at machine differences later
    message("\tRead in ", nrow(dat), " rows of data")
    message("\tDates: ", min(dat$DATE), " to ", max(dat$DATE))
    sn <- gsub("SN:\t", "", dat_raw[2], fixed = TRUE)
    message("\tInstrument serial number: ", sn)
    dat$SN <- sn
    # Parse the timezone from the header and use it to make a TIMESTAMP field
    tz <- gsub("Timezone:\t", "", dat_raw[5], fixed = TRUE)
    message("\tInstrument time zone: ", tz)
    dat$TIMESTAMP <- lubridate::ymd_hms(paste(dat$DATE, dat$TIME), tz = tz)
    dat$DATE <- dat$TIME <- NULL
  } else {
    warning("File read error for ", basefn)
    errors <<- errors + 1
    dat <- NULL
  }

  dat
}

dat <- lapply(data_files, read_ghg_data)
dat <- do.call("rbind", dat)

neg_CO2 <- sum(dat$CO2 < 0, na.rm = TRUE)
```

Errors: `r errors`

Total observations: `r nrow(dat)`

Observations with negative CO2: `r neg_CO2` (`r round(neg_CO2 / nrow(dat) * 100, 0)`%)

## Overall CO2 and CH4 data

Distribution of CO2 values:

```{r}
OVERALL_PLOT_N <- 300
summary(dat$CO2)
```

Plot of random `r OVERALL_PLOT_N` CO2 values (y axis 10%-90%):

```{r}
#| fig-width: 8
#| fig-height: 4

ylim <- quantile(dat$CO2, probs = c(0.1, 0.9), na.rm = TRUE)

dat_small <- dat[sample.int(nrow(dat), OVERALL_PLOT_N, replace = TRUE),]
ggplot(dat_small, aes(TIMESTAMP, CO2)) + 
  geom_point(na.rm = TRUE) +
  coord_cartesian(ylim = ylim)
```

Distribution of CH4 values:

```{r}
summary(dat$CH4)
```

Plot of random `r OVERALL_PLOT_N` CH4 values (y axis 10%-90%):

```{r}
#| fig-width: 8
#| fig-height: 4

ylim <- quantile(dat$CH4, probs = c(0.1, 0.9), na.rm = TRUE)

dat_small <- dat[sample.int(nrow(dat), 200, replace = TRUE),]
ggplot(dat_small, aes(TIMESTAMP, CH4)) + 
  geom_point(na.rm = TRUE) +
  coord_cartesian(ylim = ylim)
```

## Read in metadata

```{r}
errors <- 0

read_metadata <- function(fn) {
  basefn <- basename(fn)
  message(Sys.time(), " Processing ", basefn)
  
  metadat_raw <- try({
    read_excel(fn)
  })
  
  if(is.data.frame(metadat_raw)) {
    # What?
    metadat <- metadat_raw
  } else {
    warning("File read error for ", basefn)
    errors <<- errors + 1
    metadat <- NULL
  }

  metadat
}

metadat <- lapply(metadata_files, read_metadata)
metadat <- do.call("rbind", metadat)
```

Errors: `r errors`

Total rows: `r nrow(metadat)`

## Overall metadata

## Reproducibility

```{r}
sessionInfo()
```
